\emph{Light Exposure Behavior Assessment (LEBA)}: Development of a novel instrument to capture light exposure-related behaviours

### 2021-11-15 manuscript.Rmd
* All tables and figures are now cross referenced in the text

@@

For our analysis we used R (version 4.1.0), with several R packages. Initially, our tool had six point Likert type response format (0:Does not apply/I don't know; 1:Never, 2:Rarely; 3:Sometimes; 4:Often; 5:Always). Our purpose was to capture light exposure related behavior and these two response options: "Does not apply/I don't know" and "Never" were providing similar information. As such we decided to collapse them into one making it a 5 point Likert type response format. Necessary assumptions of EFA, including sample adequacy, normality assumptions, quality of correlation matrix, were assessed. Our data violated both the univariate and multivariate normality assumptions. Due to these violations and the ordinal nature of our response data, we used  polychoric correlation matrix [@desjardinsHandbookEducationalMeasurement2018] for the EFA. We employed principal axis (PA) as  factor extraction method with varimax rotation. PA is robust to the normality assumption violations [@watkinsStepbyStepGuideExploratory2020]. The obtained latent structure was confirmed by another factor extraction method: "the minimum residuals extraction" as well. We used a combination of factor identification method including scree plot [@cattellScreeTestNumber1966], Horn's parallel analysis [@hornRationaleTestNumber1965], minimum average partials method [@velicerDeterminingNumberComponents1976], and hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers. Additionally, to determine the simple structure, we followed the following guidelines recommended by psychometricians (i) no factors with fewer than three items (ii) no factors with a factor loading <0.3 (iii) no items with cross-loading greater than .3 across factors [@bandalosFactorAnalysisExploratory2018]. We confirmed the latent structure obtained in the EFA by conducting a "Confirmatory Factor Analysis" (CFA) using "robust weighted least square estimator" (WLSMV). We estiablished the measurement invariance of our tool across native and non-native English speakers using structural equation model framework. To assess the possible semantic overlap of our tool with the existing tools, we sought to "Semantic Scale Network" [@rosenbusch2020semantic]. Lastly, we sought "Item Response Theory" (IRT) based analysis on developing a short form of LEBA.
We also conducted psychometric analysis on non-merged response options data (Supp. Table \@ref(tab:EFATableAppB)) and rejected the latent structure obtained as the factors were less interpretable.

@@

min.cor <- abs(min(upper)) #minimum cor.coefficient of the matrix
max.cor <- abs(max(upper)) # max. correlation coefficient of the matrix

@@

Sampling adequacy was checked using Kaiser-Meyer-Olkin (KMO) measures of sampling adequacy [@kaiserIndexFactorialSimplicity1974] . The overall KMO vale for 48 items was `r printnum(KMO$MSA)` which was above the cutoff value (.50) indicating a mediocre sample [@hutchesonMultivariateSocialScientist1999].  Bartlett's test of sphericity [@bartlettNoteMultiplyingFactors1954], $\chi^2$ (`r bartlet$df`) = `r apa(bartlet$chisq, 2, T)`, p < .001 indicated the correlations between items are adequate for the EFA. However only `r apa(cor.per, 2,T)`% of the inter-item correlation coefficients were greater than .30. The inter item correlation ranged between `r apa(min.cor,2,F)` to `r apa(max.cor,2,F)`.

@@

With the initial 48 items we conducted three rounds of EFA and gradually discarded problematic items. (cross-loading items and poor factor loading (<.30) items). Finally, a five-factor EFA solution with 25 items was accepted with low RMSR = 0.08 [@brownConfirmatoryFactorAnalysis2015], all factor-loading higher than .30 and no cross-loading greater than .30. We further confirmed this five-factor latent structure by another EFA using varimax rotation with a minimum residual extraction method (Table \@ref(tab:MinResTab)). Table \@ref(tab:EFATable) displays the factor-loading (structural coefficients) and communality of the items. The absolute value of the factor-loading ranged from `r apa(min.loadings,2,F)` to `r apa(max.loadings, 2, F)` indicating strong coefficients. The commonalities ranged between `r apa(min.com,2,F)` to `r apa(max.com, 2, F)`. However, the histogram of the absolute values of non-redundant residual-correlations (Figure \@ref(fig:EFAResiduals)) showed `r apa(sumR,0, T)`% correlations greater than the absolute value of .05, indicating a possible under-factoring. [@desjardinsHandbookEducationalMeasurement2018]. Subsequently, we fitted a six-factor solution. However, a factor emerged with only one salient variable loading in the six-factor solution, thus disqualifying the six-factor solution (Table \@ref(tab:sixFacTab)).

@@

```{r mod2, include=FALSE}
modfit.Cor.one <- modindices(fit.Cor.2, sort. = TRUE) 
modfit.Cor.one[modfit.Cor.one$mi>3.84,]
```
We conducted categorical confirmatory factor analysis with robust weighted least square (WLSMV) estimator since our response data was of ordinary nature [@desjardinsHandbookEducationalMeasurement2018]. Several indices are suggested to measure model fit which can be categorized as absolute, comparative and parsimony fit indices [@brownConfirmatoryFactorAnalysis2015]. Absolute fit assess the model fit at an absolute level using indices including $\chi^2$ test statistics and the standardized root mean square (SRMR). Parsimony fit indices including the root mean square error of approximation (RMSEA) considers the number of free parameters in the model to assess the parsimony of the model. Comparative fit indices evaluate the fit of the specified model solution in relation to a more restricted baseline model restricting all covariances among the idicators as zero. Comparative fit index (CFI) and the Tucker Lewis index (TLI) are such two comparative fit indices. Commonly used Model fit guidelines [@huCutoffCriteriaFit1999; @schumacker2004beginner] includes (i) Reporting of $\chi^2$ test statistics (A non-significant test statistics is required to reflect model fit) (ii) CFI and TLI (CFI/TLI close to .95 or above/ranging between 90-95 and above) (iii) RMSEA (close to .06 or below), (iv) SRMR (close to .08 or below) to estimate the model fit. Table \@ref(tab:tabCfa) summarizes the fit indices of our fitted model. Our fitted model failed to attain an absolute fit estimated by the $\chi^2$ test. However, the $\chi^2$ test is sensitive to sample size and not recommended to be used as the sole index of absolute model fit [@brownConfirmatoryFactorAnalysis2015]. Another absolute fit index we obtained in our analysis was SRMR which does not work well with categorical data [@RN1272]. We judged the model fit based on the comparative fit indices: CFI, TLI and parsimony fit index:RMSEA. Our fitted model attained acceptable fit (CFI =`r apa(fit.1[5],2,F)`; TLI = `r apa(fit.1[6],2,F)`); RMSEA = `r apa(fit.1[7],2,F)`,[`r apa(fit.1[8],2,F)`-`r apa(fit.1[9],2,F)`, 90% CI]) with two imposed equity constrain on item pairs 32-33 [`r Hmisc::label(sem.data$item32)`;`r Hmisc::label(sem.data$item33)`] and 16-17 [`r Hmisc::label(sem.data$item16)`;`r Hmisc::label(sem.data$item17)`]. Items pair 32-33 stemed from the preference of dimming electric device's brightness before bed time and items pair 16 and 19 stemed from the preference of using blue filtering or colored glasses during the daytime. Nevertheless, SRMR value was higher than the guideline (SRMR = `r apa(fit.1[10],2,F)`). Further by allowing one pair of items (30-41) [`r Hmisc::label(sem.data$item30)`;`r Hmisc::label(sem.data$item41)`] to covary their error variance and discarding two item (item 37 & 26) for very low r-square value, our model attained best fit (CFI =`r apa(fit.2[5],2,F)`; TLI = `r apa(fit.2[6],2,F)`); RMSEA = `r apa(fit.2[7],2,F)`[`r apa(fit.2[8],2,F)`-`r apa(fit.2[9],2,F)`, 90% CI]) and SRMR value (SRMR = `r apa(fit.2[10],2,F)`) was also close to the suggestions of @huCutoffCriteriaFit1999. 
Internal consistency ordinal $\alpha$ for the five factors of LEBA were `r apa(rel.k[2,1],2,F)`, `r apa(rel.k[2,2],2,F)`, `r apa(rel.k[2,3],2,F)`, `r apa(rel.k[2,4],2,F)`, `r apa(rel.k[2,5],2,F)` respectively. We also estimated the internal consistency reliability of the total scale using Mcdonald's $\omega$ (total) coefficient which is a better reliability estimate for multidimensional constructs [@dunn2014alpha; @sijtsma2009use]. McDonald's $\omega$(total) coefficient for the total scale was `r apa(rel.k[5,6],2,F)`.

@@

Test information curve (TIC) (Figure \@ref(fig:Testinfo)) indicate the amount of information an the full-scale carry along the latent trait continuum. As we treated each factor of short-LEBA as an unidmensional construct we obtain 5 TICs (Figure \@ref(fig:Testinfo)). These information curves indicated except blue filter factor, the other factor's TICs are roughly centered on the center of the trait continuum ($\theta$). Also the amount of information changed rather steadily with the change of ($\theta$). Thus we conferred the LEBA tool (except blue filter) estimated the light exposure related behavior with precision near the center of trait continuum [@bakerBasicsItemResponse2017] which is sufficient to discriminate between latent trait measured by the each factor. The blue filter factor had a peak to the right side of the center of latent trait indicating its ability to providing information only for people who already have some preference towards using blue-filters.

>>>>

Figure \@ref(fig:FlowchartFig) summarizes the steps of our psychometric analysis. For our analysis we used R (version 4.1.0), with several R packages. Initially, our tool had six point Likert type response format (0:Does not apply/I don't know; 1:Never, 2:Rarely; 3:Sometimes; 4:Often; 5:Always). Our purpose was to capture light exposure related behavior and these two response options: "Does not apply/I don't know" and "Never" were providing similar information. As such we decided to collapse them into one making it a 5 point Likert type response format. Necessary assumptions of EFA, including sample adequacy, normality assumptions, quality of correlation matrix, were assessed. Our data violated both the univariate and multivariate normality assumptions. Due to these violations and the ordinal nature of our response data, we used  polychoric correlation matrix [@desjardinsHandbookEducationalMeasurement2018] for the EFA. We employed principal axis (PA) as  factor extraction method with varimax rotation. PA is robust to the normality assumption violations [@watkinsStepbyStepGuideExploratory2020]. The obtained latent structure was confirmed by another factor extraction method: "the minimum residuals extraction" as well. We used a combination of factor identification method including scree plot [@cattellScreeTestNumber1966], Horn's parallel analysis [@hornRationaleTestNumber1965], minimum average partials method [@velicerDeterminingNumberComponents1976], and hull method [@lorenzo-sevaHullMethodSelecting2011] to identify factor numbers. Additionally, to determine the simple structure, we followed the following guidelines recommended by psychometricians (i) no factors with fewer than three items (ii) no factors with a factor loading <0.3 (iii) no items with cross-loading greater than .3 across factors [@bandalosFactorAnalysisExploratory2018]. We confirmed the latent structure obtained in the EFA by conducting a "Confirmatory Factor Analysis" (CFA) using "robust weighted least square estimator" (WLSMV). We estiablished the measurement invariance of our tool across native and non-native English speakers using structural equation model framework. To assess the possible semantic overlap of our tool with the existing tools, we sought to "Semantic Scale Network" [@rosenbusch2020semantic]. Lastly, we sought "Item Response Theory" (IRT) based analysis on developing a short form of LEBA.
We also conducted psychometric analysis on non-merged response options data (Supp. Table \@ref(tab:EFATableAppB)) and rejected the latent structure obtained as the factors were less interpretable.

@@

@@

min.cor <- apa(min(abs(upper)),2,F) #minimum cor.coefficient of the matrix
max.cor <- apa(max(abs(upper)),2,F) # max. correlation coefficient of the matrix

@@

Sampling adequacy was checked using Kaiser-Meyer-Olkin (KMO) measures of sampling adequacy [@kaiserIndexFactorialSimplicity1974] . The overall KMO vale for 48 items was `r printnum(KMO$MSA)` which was above the cutoff value (.50) indicating a mediocre sample [@hutchesonMultivariateSocialScientist1999].  Bartlett's test of sphericity [@bartlettNoteMultiplyingFactors1954], $\chi^2$ (`r bartlet$df`) = `r apa(bartlet$chisq, 2, T)`, p < .001 indicated the correlations between items are adequate for the EFA. However only `r apa(cor.per, 2,T)`% of the inter-item correlation coefficients were greater than .30. The absolute value of inter-item correlation ranged between `r min.cor` to `r max.cor`. Figure \@ref(fig:figCor) depicts the correlation matrix.

@@

With the initial 48 items we conducted three rounds of EFA and gradually discarded problematic items. (cross-loading items and poor factor loading (<.30) items). Finally, a five-factor EFA solution with 25 items was accepted with low RMSR = 0.08 [@brownConfirmatoryFactorAnalysis2015], all factor-loading higher than .30 and no cross-loading greater than .30. We further confirmed this five-factor latent structure by another EFA using varimax rotation with a minimum residual extraction method (Table \@ref(tab:MinResTab)). Table \@ref(tab:EFATable) displays the factor-loading (structural coefficients) and communality of the items. The absolute value of the factor-loading ranged from `r apa(min.loadings,2,F)` to `r apa(max.loadings, 2, F)` indicating strong coefficients. The commonalities ranged between `r apa(min.com,2,F)` to `r apa(max.com, 2, F)`. Figure \@ref(fig:EFAplot) depicts the obtained five-five factor structure. However, the histogram of the absolute values of non-redundant residual-correlations (Figure \@ref(fig:EFAResiduals)) showed `r apa(sumR,0, T)`% correlations greater than the absolute value of .05, indicating a possible under-factoring. [@desjardinsHandbookEducationalMeasurement2018]. Subsequently, we fitted a six-factor solution. However, a factor emerged with only one salient variable loading in the six-factor solution, thus disqualifying the six-factor solution (Table \@ref(tab:sixFacTab)).

@@

```{r mod2, include=FALSE}
modfit.Cor.one <- modindices(fit.Cor.2, sort. = TRUE) 
modfit.Cor.one[modfit.Cor.one$mi>3.84,]
```
We conducted categorical confirmatory factor analysis with robust weighted least square (WLSMV) estimator since our response data was of ordinary nature [@desjardinsHandbookEducationalMeasurement2018]. Several indices are suggested to measure model fit which can be categorized as absolute, comparative and parsimony fit indices [@brownConfirmatoryFactorAnalysis2015]. Absolute fit assess the model fit at an absolute level using indices including $\chi^2$ test statistics and the standardized root mean square (SRMR). Parsimony fit indices including the root mean square error of approximation (RMSEA) considers the number of free parameters in the model to assess the parsimony of the model. Comparative fit indices evaluate the fit of the specified model solution in relation to a more restricted baseline model restricting all covariances among the idicators as zero. Comparative fit index (CFI) and the Tucker Lewis index (TLI) are such two comparative fit indices. Commonly used Model fit guidelines [@huCutoffCriteriaFit1999; @schumacker2004beginner] includes (i) Reporting of $\chi^2$ test statistics (A non-significant test statistics is required to reflect model fit) (ii) CFI and TLI (CFI/TLI close to .95 or above/ranging between 90-95 and above) (iii) RMSEA (close to .06 or below), (iv) SRMR (close to .08 or below) to estimate the model fit. Table \@ref(tab:tabCfa) summarizes the fit indices of our fitted model. Our fitted model failed to attain an absolute fit estimated by the $\chi^2$ test. However, the $\chi^2$ test is sensitive to sample size and not recommended to be used as the sole index of absolute model fit [@brownConfirmatoryFactorAnalysis2015]. Another absolute fit index we obtained in our analysis was SRMR which does not work well with categorical data [@RN1272]. We judged the model fit based on the comparative fit indices: CFI, TLI and parsimony fit index:RMSEA. Our fitted model attained acceptable fit (CFI =`r apa(fit.1[5],2,F)`; TLI = `r apa(fit.1[6],2,F)`); RMSEA = `r apa(fit.1[7],2,F)`,[`r apa(fit.1[8],2,F)`-`r apa(fit.1[9],2,F)`, 90% CI]) with two imposed equity constrain on item pairs 32-33 [`r Hmisc::label(sem.data$item32)`;`r Hmisc::label(sem.data$item33)`] and 16-17 [`r Hmisc::label(sem.data$item16)`;`r Hmisc::label(sem.data$item17)`]. Items pair 32-33 stemed from the preference of dimming electric device's brightness before bed time and items pair 16 and 19 stemed from the preference of using blue filtering or colored glasses during the daytime. Nevertheless, SRMR value was higher than the guideline (SRMR = `r apa(fit.1[10],2,F)`). Further by allowing one pair of items (30-41) [`r Hmisc::label(sem.data$item30)`;`r Hmisc::label(sem.data$item41)`] to covary their error variance and discarding two item (item 37 & 26) for very low r-square value, our model attained best fit (CFI =`r apa(fit.2[5],2,F)`; TLI = `r apa(fit.2[6],2,F)`); RMSEA = `r apa(fit.2[7],2,F)`[`r apa(fit.2[8],2,F)`-`r apa(fit.2[9],2,F)`, 90% CI]) and SRMR value (SRMR = `r apa(fit.2[10],2,F)`) was also close to the suggestions of @huCutoffCriteriaFit1999. Internal consistency ordinal $\alpha$ for the five factors of LEBA were `r apa(rel.k[2,1],2,F)`, `r apa(rel.k[2,2],2,F)`, `r apa(rel.k[2,3],2,F)`, `r apa(rel.k[2,4],2,F)`, `r apa(rel.k[2,5],2,F)` respectively. We also estimated the internal consistency reliability of the total scale using Mcdonald's $\omega$ (total) coefficient which is a better reliability estimate for multidimensional constructs [@dunn2014alpha; @sijtsma2009use]. McDonald's $\omega$(total) coefficient for the total scale was `r apa(rel.k[5,6],2,F)`. Figure \@ref(fig:figcfa) depicts the obtained CFA structure.

@@

Test information curve (TIC) (Figure \@ref(fig:Testinfo)) indicate the amount of information an the full-scale carry along the latent trait continuum. As we treated each factor of short-LEBA as an unidmensional construct we obtain 5 TICs (Figure \@ref(fig:Testinfo)). These information curves indicated except the first and fifth factors, the other three factor's TICs are roughly centered on the center of the trait continuum ($\theta$).The first and fifth factor had a peak to the right side of the center of latent trait.Thus we conferred the LEBA tool estimated the light exposure related behavior with precision near the center of trait continuum for 2nd, 3rd and 4th factors and near the right side of the center of trait continuum for 1st and 5th factors [@bakerBasicsItemResponse2017].
